Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-25 NEUROC/sec4-experiments.tex
section4.3增加了Memory Usage of Inference, 增加了Section4.4.4 Performance Bottleneck in Inference

% added by wangyunpan, 2020-12-22 ----start----
\paragraph{Inference Stage Memory Usage Analysis}

% 不同点: phase
\begin{figure}[H]
    \centering
    \includegraphics[height=5cm]{figs/experiments/exp_inference_full_memory_usage_stage_amp.pdf}
    \caption{Memory usage of each phase during the GNN inference. Dataset: \texttt{amp}.}
    \label{fig:exp_inference_full_memory_usage_stage_amp}
\end{figure}

% 不同点： 膨胀比例
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_full_memory_expansion_ratio.pdf}
    \caption{Memory expansion ratios of typical GNNs.}
    \label{fig:exp_inference_full_memory_expansion_ratio}
\end{figure}

% 不同点: input feature dims
\begin{figure}[H]
    \centering
    \includegraphics[height=5cm]{figs/experiments/exp_inference_full_memory_expansion_ratio_input_feature_dimension_com-amazon_expansion_ratio.pdf}
    \caption{Memory expansion ratio under different dimensions of input feature vectors during inference. Dataset: \texttt{cam}.}
    \label{fig:exp_inference_full_memory_expension_ratio_input_feature_dimension}
\end{figure}

For the inference stage, we did the same experiment as above. The experimental results were as follows:
%
\figurename~\ref{fig:exp_inference_full_memory_usage_stage_amp} shows the peak memory usage of each phase during inference stage.  
%
Compared with the results of training stage in \figurename~\ref{fig:exp_memory_usage_stage_amp}, the peak memory usage of inference stage was significantly lower.
%
This was because the inference stage didn't require backward propagation and the peak memory was only determined by the temporary variables generated during the computation.
%

\figurename~\ref{fig:exp_inference_full_memory_expansion_ratio} compares MER of different GNNs on inference stage. 
%
We found that The performance of GNNs on all datasets was consistent with the training stage in \figurename~\ref{fig:exp_memory_expansion_ratio}, but the corresponding values were lower.
%

%
% added by wangyunpan, 2020-12-22 ----end----

@@

>>>>

\subsubsection{Memory Usage of Inference}

Figure~\ref{fig:exp_inference_memory_expansion_ratio} shows the memory expansion ratios of typical GNNs during inference.
%
Since no intermediate results had to be cached during inference of GGNN, GAT, and GaAN,
the memory expansion ratios of inference were much less than training for the three GNNs (\figurename~\ref{fig:exp_memory_expansion_ratio}).
%
The MERs of inference were 45\% to 83\% (GGNN), 52\% to 61\% (GAT), and 37\% to 69\% (GaAN) of training.
%
However, the MERs were still high, disallowing inferencing with big graphs.
%
To handle big graphs, sample-based inference was necessary.

\begin{figure}[H]
   \centering
   \subfloat{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_full_memory_expansion_ratio.pdf}}
   \caption{Memory expansion ratios of typical GNNs in Inference.}
   \label{fig:exp_inference_memory_expansion_ratio}
\end{figure}

@@

\subsubsection{Performance Bottlenecks in Inference}

Sample-based inference was an effective technique to handle big graphs.
%
Taking the node classification task as an example, to predict labels for a given set of vertices $\mathcal{V}_{predict}$, the inference sampler sampled a subgraph containing complete $L$-hop neighborhoods of all vertices in $\mathcal{V}_{predict}$, where $L$ was the number of GNN layers.
%
Since the real-world graphs often had small-world property, the size of the sampled subgraph increased quickly as the number of vertices in $\mathcal{V}_{predict}$ (i.e., batch size) increased.
%
Figure~\ref{fig:sampled_graph_size} (Figure X in the revised manuscript) shows how the relative batch size ($\frac{|\mathcal{V}_{predict}|}{|\mathcal{V}|}$) affected the average degree and the number of edges of the sampled subgraphs during inference.
%
When the relative batch size was 10\%, the number of edges in the sampled subgraphs were close to the number of edges in the whole graph in most datasets.
%
In order to limit the memory usage during inference, the batch size used should be very small.

\begin{figure}[H]
   \centering
   \subfloat[Average degree]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_sampling_graph_info_inference_sampler_avg_degrees.pdf}}
   \subfloat[Number of edges]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_sampling_graph_info_inference_sampler_edges.pdf}}

   \caption{Characteristics of the sampled subgraphs produced by the inference sampler under different relative batch sizes.}
   \label{fig:sampled_graph_size}
\end{figure}

However, the overheads brought by sampling and data transfering from CPU to GPU became obvious when the batch size was small.
%
Figure~\ref{fig:time_breakdown_of_inference_sampler} shows the time breakdowns of sample-based inference on different datasets.
%
On the \texttt{amp} and \texttt{cam} datasets, the sampling time even accounted for near half of the total inference time.
%
The results indicated that the current implementation of inference sampler in PyG was also inefficient.
%
Improving its efficiency could significantly reduce the inference time.

\begin{figure}[H]
   \centering
   \subfloat[Batch Size = 1024]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_sampling_fix_batch_size_train_time_stack_1024.pdf}}
   \subfloat[Batch Size = 2048]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_sampling_fix_batch_size_train_time_stack_2048.pdf}}
   \caption{Inference time per batch breakdown under different batch sizes.}
   \label{fig:time_breakdown_of_inference_sampler}
\end{figure}
